Hazelnut: A Bidirectionally Typed \\ Structure Editor Calculus

### 2016-10-06 example.tex
adding text so people aren't confused by lack of zipper rules, close #2

@@

\begin{figure}[t!]
\begin{center}
$\arraycolsep=4px
\begin{array}{|r||l|l||l|l|}
\hline
\# & \textbf{Z-Expression} &
\textbf{Next Action} & \textbf{Rule}
\\
\hline
1 &
\zwsel{\hhole{}} &
\aConstruct{\flam{x}} &
\text{(\ref{r:conelamhole})}
\\ 2 &
\hlam{x}{\hhole{}} : \tarr{\zwsel{\hhole{}}}{\hhole{}} &
\aConstruct{\fnum{}} &
\text{(\ref{r:contnum})}
\\ 3 &
\hlam{x}{\hhole{}} : \tarr{\zwsel{\tnum{}}}{\hhole{}} &
\aMove{\dParent} &
\text{({\ref{rule:move-parent-arr-left}})}
\\ 4 &
\hlam{x}{\hhole{}} : \zwsel{\tarr{\tnum{}}{\hhole{}}} &
\aMove{\dChildn{2}} &
\text{({\ref{rule:move-arr-c2}})}
\\ 5 &
\hlam{x}{\hhole{}} : \tarr{\tnum}{\zwsel{\hhole{}}}
&
\aConstruct{\fnum{}} &
\text{(\ref{r:contnum})}
\\ 6 &
\hlam{x}{\hhole{}} : \tarr{\tnum{}}{\zwsel{\tnum{}}} &
\aMove{\dParent{}} &
\text{(\ref{rule:move-parent-arr-right})}
\\ 7 &
\hlam{x}{\hhole{}} : \zwsel{\tarr{\tnum{}}{\tnum{}}}
&
\aMove{\dParent{}} &
\text{(\ref{rule:move-parent-asc-right})}
\\ 8 &
\zwsel{\hlam{x}{\hhole{}} : \tarr{\tnum}{\tnum}} &
\aMove{\dChildn{1}} &
\text{(\ref{r:movefirstchild-asc})}
\\ 9 &
\zwsel{\hlam{x}{\hhole{}}} : \tarr{\tnum{}}{\tnum{}} &
\aMove{\dChildn{1}} &
\text{(\ref{r:movefirstchild-lam})}
\\ 10 &
\hlam{x}{\zwsel{\hhole{}}} : \tarr{\tnum{}}{\tnum{}} &
\aConstruct{\fvar{x}} &
\text{(\ref{r:conevar})}
\\ 11 &
\hlam{x}{\zwsel{{x}}} : \tarr{\tnum{}}{\tnum{}} &
{\aConstruct{\fplus}}
&
\text{(\ref{rule:construct-plus-compat})}
\\ 12 &
\hlam{x}{\hadd{x}{\zwsel{\hhole{}}}} : \tarr{\tnum{}}{\tnum{}} &
\aConstruct{\fnumlit{1}} &
\text{(\ref{r:conenumnum})}
\\ 13 &
\hlam{x}{\hadd{x}{\zwsel{\hnum{1}}}} : \tarr{\tnum}{\tnum} &
\textrm{---} &
{\textrm{---}}
\\ \hline
\end{array}
$\end{center}\vspace{-6px}
\caption{Constructing an increment function in Hazelnut.}
\label{fig:first-example}
\end{figure}

@@

\begin{figure}[t!]
\begin{center}
\colorbox{light-gray}{\hspace{53px} now assume $incr : \tarr{\tnum}{\tnum}$ \hspace{54px}}
$\arraycolsep=4px
\begin{array}{|r||l|l||l|l|}
\hline
\# & \textbf{Z-Expression} &
\textbf{Next Action} & \textbf{Rule}
\\
\hline
14 &
\zwsel{\hhole{}} &
\aConstruct{\fvar{incr}} \hphantom{~\,\,~~}&
\text{(\ref{r:conevar})}
\\ 15 &
\zwsel{incr} &
\aConstruct{\fap} &
\text{(\ref{r:coneapfn})}
\\ 16 &
incr(\zwsel{\hhole{}}) &
\aConstruct{\fvar{incr}} &
\text{(\ref{r:conevar2})}
\\ 17 &
incr(\hhole{\zwsel{incr}}) &
\aConstruct{\fap} &
\text{(\ref{r:coneapfn})}
\\ 18 &
incr(\hhole{incr(\zwsel{\hhole{}})}) \hphantom{~~~~} &
\aConstruct{\fnumlit{3}} &
\text{(\ref{r:conenumnum})}
\\ 19 &
incr(\hhole{incr(\zwsel{\hnum{3}})}) &
\aMove{\dParent}&
\text{(\ref{r:moveparent-ap2})}
\\ 20 &
incr(\hhole{\zwsel{incr(\hnum{3})}}) &
\aMove{\dParent} &
\text{(\ref{r:moveparent-hole})}
\\ 21 &
incr(\zwsel{\hhole{incr(\hnum{3})}})&
\aFinish &
\text{(\ref{r:finishana})}
\\ 22 &
incr(\zwsel{incr(\hnum{3})}) &
\textrm{---} &
{\textrm{---}}
\\ \hline
\end{array}
$\end{center}\vspace{-6px}
\caption{Appyling the increment function.}
\label{fig:second-example}
\end{figure}

>>>>

For concision, the right column in both Fig. \ref{ex1} and Fig \ref{ex2}
lists only the rules that describe the edit \emph{at the cursor}. This
omits both the rules needed to work down the tree structure of the
{Z-expression} to the cursor and the rules to propagate the change back
up. The missing rules are described in detail in
Sec. \ref{sec:zipper-cases}, and these examples are included in full in the
Agda mechanization described in Sec. \ref{sec:mech}.

@@

\begin{figure}[t!]
  \label{ex1}
\begin{center}
$\arraycolsep=4px
\begin{array}{|r||l|l||l|l|}
\hline
\# & \textbf{Z-Expression} &
\textbf{Next Action} & \textbf{Rule}
\\
\hline
1 &
\zwsel{\hhole{}} &
\aConstruct{\flam{x}} &
\text{(\ref{r:conelamhole})}
\\ 2 &
\hlam{x}{\hhole{}} : \tarr{\zwsel{\hhole{}}}{\hhole{}} &
\aConstruct{\fnum{}} &
\text{(\ref{r:contnum})}
\\ 3 &
\hlam{x}{\hhole{}} : \tarr{\zwsel{\tnum{}}}{\hhole{}} &
\aMove{\dParent} &
\text{({\ref{rule:move-parent-arr-left}})}
\\ 4 &
\hlam{x}{\hhole{}} : \zwsel{\tarr{\tnum{}}{\hhole{}}} &
\aMove{\dChildn{2}} &
\text{({\ref{rule:move-arr-c2}})}
\\ 5 &
\hlam{x}{\hhole{}} : \tarr{\tnum}{\zwsel{\hhole{}}}
&
\aConstruct{\fnum{}} &
\text{(\ref{r:contnum})}
\\ 6 &
\hlam{x}{\hhole{}} : \tarr{\tnum{}}{\zwsel{\tnum{}}} &
\aMove{\dParent{}} &
\text{(\ref{rule:move-parent-arr-right})}
\\ 7 &
\hlam{x}{\hhole{}} : \zwsel{\tarr{\tnum{}}{\tnum{}}}
&
\aMove{\dParent{}} &
\text{(\ref{rule:move-parent-asc-right})}
\\ 8 &
\zwsel{\hlam{x}{\hhole{}} : \tarr{\tnum}{\tnum}} &
\aMove{\dChildn{1}} &
\text{(\ref{r:movefirstchild-asc})}
\\ 9 &
\zwsel{\hlam{x}{\hhole{}}} : \tarr{\tnum{}}{\tnum{}} &
\aMove{\dChildn{1}} &
\text{(\ref{r:movefirstchild-lam})}
\\ 10 &
\hlam{x}{\zwsel{\hhole{}}} : \tarr{\tnum{}}{\tnum{}} &
\aConstruct{\fvar{x}} &
\text{(\ref{r:conevar})}
\\ 11 &
\hlam{x}{\zwsel{{x}}} : \tarr{\tnum{}}{\tnum{}} &
{\aConstruct{\fplus}}
&
\text{(\ref{rule:construct-plus-compat})}
\\ 12 &
\hlam{x}{\hadd{x}{\zwsel{\hhole{}}}} : \tarr{\tnum{}}{\tnum{}} &
\aConstruct{\fnumlit{1}} &
\text{(\ref{r:conenumnum})}
\\ 13 &
\hlam{x}{\hadd{x}{\zwsel{\hnum{1}}}} : \tarr{\tnum}{\tnum} &
\textrm{---} &
{\textrm{---}}
\\ \hline
\end{array}
$\end{center}\vspace{-6px}
\caption{Constructing an increment function in Hazelnut.}
\label{fig:first-example}
\end{figure}

@@

\begin{figure}[t!]
  \label{ex2}
\begin{center}
\colorbox{light-gray}{\hspace{53px} now assume $incr : \tarr{\tnum}{\tnum}$ \hspace{54px}}
$\arraycolsep=4px
\begin{array}{|r||l|l||l|l|}
\hline
\# & \textbf{Z-Expression} &
\textbf{Next Action} & \textbf{Rule}
\\
\hline
14 &
\zwsel{\hhole{}} &
\aConstruct{\fvar{incr}} \hphantom{~\,\,~~}&
\text{(\ref{r:conevar})}
\\ 15 &
\zwsel{incr} &
\aConstruct{\fap} &
\text{(\ref{r:coneapfn})}
\\ 16 &
incr(\zwsel{\hhole{}}) &
\aConstruct{\fvar{incr}} &
\text{(\ref{r:conevar2})}
\\ 17 &
incr(\hhole{\zwsel{incr}}) &
\aConstruct{\fap} &
\text{(\ref{r:coneapfn})}
\\ 18 &
incr(\hhole{incr(\zwsel{\hhole{}})}) \hphantom{~~~~} &
\aConstruct{\fnumlit{3}} &
\text{(\ref{r:conenumnum})}
\\ 19 &
incr(\hhole{incr(\zwsel{\hnum{3}})}) &
\aMove{\dParent}&
\text{(\ref{r:moveparent-ap2})}
\\ 20 &
incr(\hhole{\zwsel{incr(\hnum{3})}}) &
\aMove{\dParent} &
\text{(\ref{r:moveparent-hole})}
\\ 21 &
incr(\zwsel{\hhole{incr(\hnum{3})}})&
\aFinish &
\text{(\ref{r:finishana})}
\\ 22 &
incr(\zwsel{incr(\hnum{3})}) &
\textrm{---} &
{\textrm{---}}
\\ \hline
\end{array}
$\end{center}\vspace{-6px}
\caption{Appyling the increment function.}
\label{fig:second-example}
\end{figure}
